Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-05-10 Ben_Manuscripts/hdphmm/Supporting_Information.tex
Updates to methods and clustering

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 	
  	\item Convert $x$ and $y$ coordinates to radial coordinates, $r$, then fix the state sequence
  	and use the inference component of the IHMM in order to estimate the parameters in terms
  	of $r$ and $z$ (see Figure~\ref{fig:rz_hmm}). 	 	
  	\item Cluster the VAR parameters from the states found in all 24 trajectories. Reassign the state
  	sequence so that segments which belong to the same cluster are labeled the same across all solute
  	trajectories.
  	\item Subtract the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  	
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{rz_hmm.pdf}
		\caption{}\label{fig:rz_hmm}
	\end{subfigure}  
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{zeroed_hmm.pdf}
		\caption{}\label{fig:zeroed_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}

>>>>

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 		
  	\item Cluster the VAR parameters from the states found in all 24 trajectories (See 
  	Section~\ref{method:clustering} of the main text). Reassign the state sequence so that segments
  	which belong to the same cluster are labeled the same across all solute trajectories.
  	\item Zero the trajectories. First, zero out the $y$ dimension by rotating each segment of 
  	the trajectory, as partitioned before clustering, about the $z$ axis, in order to align the 
  	mean $xy$ vector with the $x$ axis (see Figure~\ref{fig:y_zeroed}). Then, subtract the mean in
  	$x$ and $z$ (Figure~\ref{fig:xyz_zeroed}.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_clustered_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.6\textwidth}
	\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  
		\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{y_zeroed.pdf}
		\caption{}\label{fig:y_zeroed}
	\end{subfigure}  	
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{xyz_zeroed.pdf}
		\caption{}\label{fig:xyz_zeroed}
	\end{subfigure}  		
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{zeroed_clustered_hmm.pdf}
		\caption{}\label{fig:zeroed_clustered_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}
